Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-03-16 Paper/Manuscript.tex
Updated introduction

@@

Active acoustic remote sensing is an essential method for marine ecosystem research and monitoring. Echosounders provide unique information on marine organisms on a wide range of spatial and temporal scales, and on ecosystem components from plankton to large predators.

Management of fisheries resources have successfully employed quantitative single-beam echosounders for abundance estimates for decades, obtained through spatially extensive surveys mapping the distribution and abundance of species of fishes (Simmonds and MacLennan, 2005). Echosounders are increasingly used for the study of not only the distribution and abundance of commercially important fishes, but also the distribution of other trophic levels, and interactions among ecosystem components and drivers at the level of population, foodweb and ecosystem (Godø et al., 2014).

Single-beam echosounders have provided important insight in studies of predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).

Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz) can help in the categorizing the backscatter into species or functional groups. This is termed the multi-frequency approach (Holliday and Pieper, 1980; Korneliussen and Ona, 2002).

The application of acoustic pulses with a wide and continuous frequency range (broadband pulses) to applications in ecology is an obvious enhancement. Broadband pulses provide better frequency resolution and coverage than multi-frequency systems, i.e., a continuous frequency coverage over a wide frequency band, and with appropriate processing can provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000). These benefits can lead to improved backscatter categorization (R. Korneliussen, 2018; R. J. Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning (see e.g. Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).

@@

The increasing availability of commercial systems specifically for ecosystem acoustics will vastly increase the use of broadband pulses in aquatic ecology, as these systems are more user-friendly  and with capabilities which ease the transition from existing narrowband scientific echosounders (e.g. quantitative data output and a familiar data format supported by the major software suites for marine ecosystem acoustics).

The purpose of this paper is to present a systematic and comprehensive explanation of how to derive quantitative broadband data from recorded broadband signals with associated code, and to provide a code base for further development and research in broadband acoustic ecology.

@@

\subsection{Accompanying code}
The code accompanying this paper is written in Python and is available on GitHub. All processing steps with respective figures in the paper can be reproduced by running main script (Main.py). Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.

@@

\section{Discussion}
Obtaining quantitative broadband data is more complicated than for narrowband echosounders, due in part to the need to account for frequency dependence in most variables and parameters, and the need for an increased understanding of digital signal processing techniques. For example, the \ek{} currently provides complex demodulated \citep{hasan1983} voltages for each transducer sector rather than derived quantities (e.g., the envelope of a pulse compressed signal and the frequency response of single targets and volumes) - these outputs must instead be calculated as proposed in this paper. This was an implementation decision made to allow for more flexible use of the echosounder data, and to ease the development of new processing methodologies. This decision has several disadvantages, such as the significantly increased data quantity and markedly higher amount of computation required to simply display an echogram. These can be ameliorated to some degree by processing the data into a more directly useful form before storage. The advantages of having access to unprocessed data were considered to out-weigh these disadvantages due to the potential benefits of more sophisticated uses of the acoustic data, especially for a tool that is only beginning to be applied to the field of fisheries acosutics.

The methodology to process broadband data has been presented in a general form in the previous sections without any accord to engineering limits. However, any physical implementation introduces operational constraints. As an example, a system designed for shipboard installations with ample access to electrical and computer processing power and large data storage will typically have different operational contraints compared to a system intended for autonomous platforms where electrical power and computing resources can be severely limited. In addition, the transducer is a significant constraint on the operational parameters of an echosounder and is usually the main determinant of the usable transceiver operating parameters (such as transmit bandwidth, maximum transmit power, pulse duration, ping rate, etc).

The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years. For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}. The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

>>>>

\author{Geir Pedersen}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

Active acoustics is an efficient method for remote sensing of marine ecosystems. These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km (Makris) to small scale behaviour patterns (refs). These methods are also a crucial component of fisheries resource management. Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperature sonar, omnidirectional sonars, acoustic imaging sonars as well as echosounders (refs).

Single- beam echosounders have extensively been used in fisheries management and ecosystem monitoring. Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful. Single-beam echosounders have provided important information for fisheries resource management (Simmonds and MacLennan, 2005) as well as insight in ecological processes for the aquatic ecosystem (Godø 2014).

Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz). The method has been used to  study predator-prey interactions (Benoit-Bird and Au, 2003; Friedlaender et al., 2009; Veit et al., 1993), interactions between marine organisms and the physical environment (Crawford et al., 2012; Freeman et al., 2004; Georgiadis et al., 2009; Ludsin et al., 2009; Godø...), abundance and distribution of prey and predators (Thompson et al., 1991) and non-fish species or group of species (Brierley et al., 2001), and the individual or group behavior of marine organisms (Axelsen et al., 2000; Benoit-Bird et al., 2001; Kaartvedt et al., 2009).

Recent echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems, i.e. a continuous frequency coverage over a wide frequency band. provide significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses (Chu and Stanton, 1998; Ehrenberg and Torkelson, 2000). These benefits can lead to improved backscatter categorization (R. Korneliussen, 2018; R. J. Korneliussen et al., 2016; Martin Traykovski et al., 1998; Stanton et al., 2012) and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning 
(skriv om: e.g. Lavery et al., 2017; Basset et al., 2017; 2020; Skaret et al., 2020; Kubilius et al., 2020; Benoit-Bird and Waluk, 2020; Hasegawa et al., 2021; Gugele et al., 2021; Cotter et al., 2021; Blaunet et al., 2021; Agersted et al., 2021; Khodabandeloo et al., 2021a;2021b).

@@

To tap the full potential of broadband echosounder data, further methods for processing such data must be developed. Examples include adapting pulses to improve target separation as well as further parameters from the individual targets. Different send pulses can be envisioned as well as various methods for acoustic target classification. The objective of this paper is to present a systematic and comprehensive explanation of how to process calibrated broadband echosounder data. By making associated code available, we provide a starting point for further developing signal processing methods for applying broadband acoustic to aquatic ecology.

@@

\subsection{Accompanying code}
The code accompanying this paper is written in Python and is available as supplementary materials. All processing steps with respective figures in the paper can be reproduced by running main script (Main.py). Without loss of generality, we use the Simrad EK80 as an example since it is currently the most used broadband echosounder in the marine ecosystem acoustics field. By presenting the design goals, implementation details, and recommended procedures and processing required to obtain quantitative broadband data, the authors hope to encourage and facilitate the realistic use of broadband signals in marine ecosystem acoustics.

@@

\section{Discussion}
The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years. For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassettBroadbandEchosounderMeasurements2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret2020} and to detect objects adjacent to boundaries \citep{lavery2017}. The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

Modern machine learning techniques are well suited for further processing of broadband echosounder data. Deep learning algorithms, including deep clustering, provide powerful methods to tease out patterns from high dimensional and complex data. By providing similar step sizes across frequencies it is possible to adapt this code to for an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.
